---
title: "732A91 Lab 2"
author: "Fanny Karelius (fanka300), Milda Poceviciute (milpo192)"
date: "15 May 2018"
output: html_document
---

#Question 1: Poisson regression - the MCMC way

The $y_i$th ($i=1,...,n$) count of for the $i$th observation in the sample follows the following regression model:
$$y_i|\beta \sim Poisson[exp(x_i^{T}\beta)]$$
where $x_i$ is a $p$-dimensional vector of covariates.


The response variable is nBids and records the number of bids in each auction. The remaining variables are features/covariates (x): 
Const (for the intercept) 
PowerSeller (is the seller selling large volumes on eBay?) 
VerifyID (is the seller veri???ed by eBay?) 
Sealed (was the coin sold sealed in never opened envelope?) 
MinBlem (did the coin have a minor defect?) 
MajBlem (a major defect?) 
LargNeg (did the seller get a lot of negative feedback from customers?) 
LogBook (logarithm of the coins book value according to expert sellers. Standardized) 
MinBidShare (a variable that measures ratio of the minimum selling price (starting price) to the book value. Standardized).

##a)

```{r}
data <- read.table("eBayNumberOfBidderdata.dat", header = TRUE)
c <- c(1,3:ncol(data))
data_noC <- data[,c]
glm_fit <- glm(nBids~., family=poisson, data_noC)
summary(glm_fit)
```

##b)

##c)

##d)



#Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```